---
title: "RWorksheet_Alpanghe#4c"
author: "Jersey Gabriel E. Alpanghe"
date: "2024-11-11"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1. Use the dataset mpg
A data frame with 234 rows and 11 variables:

• Download and open the mpg file. Upload it to your OWN environment
a. Show your solutions on how to import a csv file into the environment.
```{r}
mpg <- read.csv("mpg.csv")
```
b. Which variables from mpg dataset are categorical?

Categorical variables in mpg: manufacturer, model, trans, drv, fl, class.

c. Which are continuous variables?

Continuous variables in mpg: displ, year, cyl, cty, hwy.

2. Which manufacturer has the most models in this data set? Which model has the most variations?
Show your answer.

a. Group the manufacturers and find the unique models. Show your codes and result.

```{r}
library(dplyr)

manufacturer_count <- mpg %>%
  group_by(manufacturer) %>%
  summarise(num_models = n_distinct(model)) %>%
  arrange(desc(num_models))

model_variations <- mpg %>%
  group_by(model) %>%
  summarise(num_variations = n()) %>%
  arrange(desc(num_variations))

print(manufacturer_count)
print(model_variations)

```

b. Graph the result by using plot() and ggplot(). Write the codes and its result.

```{r}
library(ggplot2)
manufacturer_count$manufacturer <- factor(manufacturer_count$manufacturer)

plot(as.numeric(manufacturer_count$manufacturer), manufacturer_count$num_models, 
     type = "b", pch = 19, col = "blue",
     main = "Number of Models by Manufacturer", 
     xlab = "Manufacturer", ylab = "Number of Models", 
     xaxt = "n")

axis(1, at = 1:length(manufacturer_count$manufacturer), 
     labels = levels(manufacturer_count$manufacturer), las = 2, cex.axis = 0.8)

```

2. Same dataset will be used. You are going to show the relationship of the modeland the manufacturer.

a. What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show?

b. For you, is it useful? If not, how could you modify the data to make it more informative?

3. Plot the model and the year using ggplot(). Use only the top 20 observations. Write the codes and
its results.

```{r}
# Top 20 observations by model and year
top_20 <- mpg %>% slice(1:20)

ggplot(top_20, aes(x = model, y = year)) +
  geom_point() +
  labs(title = "Model vs. Year for Top 20 Observations", x = "Model", y = "Year") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

4. Using the pipe (%>%), group the model and get the number of cars per model. Show codes and its result

a. Plot using geom_bar() using the top 20 observations only. The graphs shoudl have a title, labels and colors. Show code and results.

```{r}
car_count <- mpg %>%
  group_by(model) %>%
  summarise(num_cars = n()) %>%
  arrange(desc(num_cars))

print(car_count)

```
b. Plot using the geom_bar() + coord_flip() just like what is shown below. Show codes and its result.

```{r}
# Basic bar plot
ggplot(car_count %>% slice(1:20), aes(x = reorder(model, -num_cars), y = num_cars)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Number of Cars per Model (Top 20)", x = "Model", y = "Number of Cars") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Bar plot with coord_flip()
ggplot(car_count %>% slice(1:20), aes(x = reorder(model, -num_cars), y = num_cars)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Number of Cars per Model (Top 20)", x = "Model", y = "Number of Cars") +
  coord_flip()

```

5. Plot the relationship between cyl - number of cylinders and displ - engine displacement using geom_point with aesthetic color = engine displacement. Title should be “Relationship between No. of Cylinders and Engine Displacement”.

```{r}
ggplot(mpg, aes(x = cyl, y = displ, color = displ)) +
  geom_point() +
  labs(title = "Relationship between No. of Cylinders and Engine Displacement",
       x = "Number of Cylinders", y = "Engine Displacement")

```
a. How would you describe its relationship? Show the codes and its result.

6. Plot the relationship between displ (engine displacement) and hwy(highway miles per gallon). Mapped
it with a continuous variable you have identified in #1-c. What is its result? Why it produced such
output?

```{r}
ggplot(mpg, aes(x = displ, y = hwy, color = cty)) +
  geom_point() +
  labs(title = "Relationship between Engine Displacement and Highway MPG",
       x = "Engine Displacement", y = "Highway MPG")

```

6. Import the traffic.csv onto your R environment.

```{r}
traffic <- read.csv("traffic.csv")
```
a. How many numbers of observation does it have? What are the variables of the traffic dataset the Show
your answer.
```{r}
print(dim(traffic))  
print(names(traffic))
```
b. subset the traffic dataset into junctions. What is the R codes and its output?

```{r}
junction_traffic <- traffic %>% filter(Junction == "junction")
print(junction_traffic)

```

c. Plot each junction in a using geom_line(). Show your solution and output.

```{r}
ggplot(traffic, aes(x = DateTime, y = Vehicles, color = Junction)) +
  geom_line() +
  labs(title = "Traffic Counts by Junction", x = "Time", y = "Traffic Count")

```

7. From alexa_file.xlsx, import it to your environment
a. How many observations does alexa_file has? What about the number of columns? Show your solution
and answer.
```{r}
library(readxl)
alexa_data <- read_excel("alexa_file.xlsx")
print(dim(alexa_data))
```
b. group the variations and get the total of each variations. Use dplyr package. Show solution and
answer.
```{r}
alexa_variations <- alexa_data %>%
  group_by(variation) %>%
  summarise(total = n())
print(alexa_variations)

```
c. Plot the variations using the ggplot() function. What did you observe? Complete the details of the
graph. Show solution and answer.

```{r}
ggplot(alexa_variations, aes(x = variation, y = total)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  labs(title = "Total Count of Each Variation", x = "Variation", y = "Total Count")

```

d. Plot a geom_line() with the date and the number of verified reviews. Complete the details of the
graphs. Show your answer and solution.

```{r}
ggplot(alexa_data, aes(x = date, y = verified_reviews)) +
  geom_line() +
  labs(title = "Verified Reviews Over Time", x = "Date", y = "Verified Reviews")

```

e. Get the relationship of variations and ratings. Which variations got the most highest in rating? Plot a
graph to show its relationship. Show your solution and answer.

```{r}
variation_ratings <- alexa_data %>%
  group_by(variation) %>%
  summarise(avg_rating = mean(rating))

ggplot(variation_ratings, aes(x = reorder(variation, -avg_rating), y = avg_rating)) +
  geom_bar(stat = "identity", fill = "coral") +
  labs(title = "Average Rating by Variation", x = "Variation", y = "Average Rating") +
  coord_flip()

```