---
title: "RWorksheet_Alpanghe#4b"
author: "Jersey Gabriel E. Alpanghe"
date: "2024-10-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Using the for loop, create an R script that will display a 5x5 matrix as shown in Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.

```{r}
vectorA <- c(1,2,3,4,5)
matrixB <- matrix(0, nrow = 5, ncol = 5)

for (i in 1:5){
  for (j in 1:5){
    matrixB[i,j] <- abs(i - j) 
  }
}

print(matrixB)
```

2. Print the string "*" using for() function. The output should be the same as shown in Figure 2.
```{r}
for (i in 1:5) {
  cat(rep("*", i), "\n")
}
```

3. Get an input from the user to print the Fibonacci sequence starting from the 1st input up to 500. Use repeat and break statements. Write the R Scripts and its output.
```{r}
# Get the starting number from the user
start_number <- readline("Enter the starting number for the Fibonacci sequence: ")

# Convert the input to numeric
start_number <- as.numeric(start_number)

# Check if the input is valid
if (is.na(start_number) || start_number < 1) {
  cat("Invalid input. Please enter a positive integer.\n")
} else {
  a <- 0
  b <- 1
  
  cat(a, " ")
    repeat {
    c <- a + b
        cat(c, " ")
        a <- b
    b <- c
    
    if (c > 500) {
      break
    }
  }
}
```

4. Import the dataset as shown in Figure 1 you have created previously.

a. What is the R script for importing an excel or a csv file? Display the first 6 rows of the dataset? Show your codes and its result

```{r}
shoes <- read.csv("shoes.csv")
head(shoes)
```

b. Create a subset for gender(female and male). How many observations are there in Male? How about in Female? Write the R scripts and its output.

```{r}
male_data <- subset(shoes, Gender == "M")
male_count <- nrow(male_data)

female_data <- subset(shoes, Gender == "F")
female_count <- nrow(female_data)

cat("Number of Males: ", male_count, "\n")
cat("Number of Females: ", female_count, "\n")

```

```{r}

gender_counts <- table(shoes$Gender)
num_male <- nrow(shoes, Gender == "M")
num_female <- nrow(shoes, Gender == "F")

gender_counts <- c(num_male, num_female)

gender_labels <- c("Male", "Female")
colors <- c("skyblue", "pink")

barplot(
  gender_counts,
  names.arg = gender_labels,
  col = colors,
  main = "Number of Males and Females in Household Data",
  xlab = "Gender",
  ylab = "Count",
  ylim = c(0, max(gender_counts) + 2)
)

legend("topright", legend = gender_labels, fill = colors)

```

5. The monthly income of Dela Cruz family was spent on the following:

a. Create a piechart that will include labels in percentage.Add some colors and title of the chart. Write the R scripts and show its output.

```{r}
# Define the data
expenses <- c(Food = 60, Electricity = 10, Savings = 5, Miscellaneous = 25)

# Calculate percentages for labels
percent_labels <- paste0(names(expenses), ": ", round(expenses / sum(expenses) * 100, 1), "%")

# Create the pie chart
pie(
  expenses,
  labels = percent_labels,
  main = "Dela Cruz Family Monthly Expenses",
  col = c("darkgreen", "blue", "pink", "red") # Add colors for each category
)

```

6. Use the iris dataset.
data(iris)

a. Check for the structure of the dataset using the str() function. Describe what you have seen in the output.
```{r}
data(iris)

str(iris)
```



b. Create an R object that will contain the mean of the sepal.length,
sepal.width,petal.length,and petal.width. What is the R script and its result?

```{r}
mean_values <- colMeans(iris[, 1:4])

mean_values

```

c. Create a pie chart for the Species distribution. Add title, legends, and colors. Write the R script and its result.

```{r}
species <- table(iris$Species)

pie(
  species, main = "Species Distribution (Iris)",
  col = c("lightblue", "lightgreen", "lightyellow"), labels = paste(names(species), ":", species), legend.text = names(species)
)
```

d. Subset the species into setosa, versicolor, and virginica. Write the R scripts and show the last six (6) rows of each species.

```{r}
setosa <- subset(iris, Species == "setosa")
versicolor <- subset(iris, Species == "versicolor")
virginica <- subset(iris, Species == "virginica")

tail(setosa)
tail(versicolor)
tail(virginica)
```

e. Create a scatterplot of the sepal.length and sepal.width using the different
species(setosa,versicolor,virginica). Add a title = “Iris Dataset”, subtitle = “Sepal width and length, labels for the x and y axis, the pch symbol and colors should be based on the species.

```{r}
# Convert Species to a factor if it isn't already
iris$Species <- as.factor(iris$Species)

# Plot scatterplot with color coding by species
plot(
  iris$Sepal.Length, iris$Sepal.Width,
  main = "Iris Dataset",
  sub = "Sepal Width and Length",
  xlab = "Sepal Length",
  ylab = "Sepal Width",
  pch = 19, # Symbol for points
  col = as.numeric(iris$Species)
)

# Add a legend for species
legend(
  "topright",
  legend = levels(iris$Species),
  col = 1:3, pch = 19
)

```


f. Interpret the result.

The different colors for each species show clear clusters, indicating that the species have distinct sepal measurements. Setosa tends to be well-separated from versicolor and virginica in terms of both Sepal.Length and Sepal.Width. versicolor and virginica, however, may overlap slightly, especially in sepal width, but still display a general trend that helps distinguish them.